---
title: "Root Sample Inventory for Sequencing"
output: html_notebook
---

```{r}
library(tidyr)
library(dplyr)
```

```{r, echo = FALSE}
##set prefix for home directory
#windows
home = 'C:/Users/Bryan/'
##for mac
#home = "~/"

```

2017 file metadata was extracted from sample codes

```{r}
Samples2017 = read.delim(file = paste(home,'Box/research/BTI/DOE_SorghumMicrobiome/AMF_PacBio/Proposal/Samples_wMetadata_2017.txt', sep =''), header = TRUE, colClasses = rep("factor"))
head(Samples2017)



```

# Check 2016 sample list
2016 metadata is in separate file, import JGI sample sheet from allyn and join with metadata file
```{r}
Samples2016 = read.delim(file = paste(home,'Box/research/BTI/DOE_SorghumMicrobiome/AMF_PacBio/Proposal/2016SamplesForPacBio.txt', sep =''), header = TRUE, colClasses = rep("factor"), sep = "\t")
head(Samples2016)

Md2016 = read.delim(file = paste(home,'Box/research/BTI/DOE_SorghumMicrobiome/AMF_PacBio/Proposal/JGI_samples_2016_031518.txt', sep =''), sep = "\t", header = TRUE, colClasses = rep("factor"))
head(Md2016)

```

* The following samples are duplicated in the Root samples 2016_2017 metadata.xlsx file
```{r}
Samples2016[duplicated(Samples2016),]$Sample.Name. 
```


* create new sample name for linking with JGI metadata sheet
```{r}
Samples2016 = Samples2016 %>% 
  mutate(tmp = gsub("R","", .$Sample.Name.)) %>%
  mutate(SampleName = gsub("-", "_", .$tmp)) %>%
  select(-tmp)
head(Samples2016)
```

## Are any samples in the list that don't have metadata?
```{r}
Mdlist = Md2016$JGI_Sample_ID %>% as.character()
Samples2016 %>% filter(!SampleName %in% Mdlist) 
```
* There are 35 samples that names do not match the metadata file
* Find an example in the metadata file
```{r}
Md2016 %>% filter(Location == "Central City, NE" & Date == "2016-07-05" & Geno == "E17" & Rep == "9" & Plot == "16")
```

* it appears that the first sample checked simply has "-" seperators instead of "_"
* For purposes of the summary table, all "-" will be substituted with "_" in the metadata file.  Alert Allyn in case the error is important

* Replace dashes with underscore and then check if there are any samples qued for pac bio that are not in the metadata file
```{r}
Md2016$JGI_Sample_ID = gsub("-", "_", Md2016$JGI_Sample_ID)
Mdlist = Md2016$JGI_Sample_ID %>% as.character()
Samples2016 %>% filter(!SampleName %in% Mdlist) 

```
* After gsub there are no samples in PacBio list from 2016 that are not in the metadata file

## Make summary table for 2016

```{r}
#get list of samples for pac bio sequencing and filter metadata table
Samples2016 = Samples2016 %>% unique
dim(Samples2016)
Slist = Samples2016$SampleName %>% as.character()
#Slist %>% length() %>% paste(., "records in pacbio sample list")
Slist %>% unique %>% length() #%>% paste("but only", ., "are unique")

PB.md = Md2016 %>% filter(JGI_Sample_ID %in% Slist)
dim(PB.md)
head(PB.md)
colnames(PB.md)
```

```{r}
#remove dashes from date field
PB.md$Date = gsub("-","",PB.md$Date)
SampleTable2016 = PB.md %>%
  mutate(DateSite = paste(substr(JGI_Sample_ID,1,2),Date,sep = '_'),
         Treatment = factor(Treat, levels = c("Full N","Low N", "WW", "WS"))) %>%
  group_by(Geno, DateSite, Treat) %>%
  summarize(n = length(Plot)) %>%
  as.data.frame() %>%
  spread(key = DateSite, value = n) %>%
  arrange(desc(CC_20160705), desc(SB_20160827), Geno)
SampleTable2016
```


*Scottsbluff samples from 20160707 are missing, except one treatment with four reps. Samples appear on sample summary sheet (and JGI metadata sheet), but not on PacBio sheet

*In the central city samples from 20160913 and 20161011 there appear to be an missing and extra rep respectively.  This appears to be a result of inaccuracies in the metadata database file rather than the pacbio file (CC0126F_E14_091316_1813, CC0127F_E17_091316_1814, CC0101F_G1_091316_1812 )

```{r}
PB.md$Date %>% head
PB.md %>% filter(Location == "Central City, NE", Date %in% c("20160913", "20161011")  & Treat == "Full N") %>% arrange(Geno, Date)
```

* Two samples missing from E13 at scotts bluff in the well-watered treatment

# 2017 sample summary
```{r}
Samples2017 %>%
  group_by(Line, Date, Site, Treatment) %>%
  summarize(n = length(Plot)) %>%
  head
```

```{r}
SampleTable2017 = Samples2017 %>%
  mutate(DateSite = paste(Site,Date,sep = '_'),
         Treatment = factor(Treatment, levels = c("F","L", "W", "D"))) %>%
  group_by(Line, DateSite, Treatment) %>%
  summarize(n = length(Plot)) %>%
  as.data.frame() %>%
  spread(key = DateSite, value = n) %>%
  arrange(desc(CC_062817), desc(SB_070717), Line)
SampleTable2017
```

